Towards a Practical Face Recognition System: \\ Robust Alignment and Illumination by Sparse Representation

### 2011-05-08 release_version/main.tex
Typesetting the same up to page 10

\begin{figure}
\centering
\includegraphics{Figure_2}
\caption{{\bf Comparing alignment of a subject wearing sunglasses by
$\ell^1$ and $\ell^2$ minimization.}
{\bf Top:} alignment result of minimizing $\|\e\|_1$; {\bf Bottom:}
result of minimizing $\|\e\|_2$. (a) {\em Green (dotted):} Initial face boundary
given by the face detector, {\em Red (solid):} Alignment result shown on the same
face; (b) warped testing image using the estimated transformation $\y_0$;
(c) reconstructed face $A_i\x$ using the training; (d) image of error $\e$. }\label{fig:L1-L2-align}
\vspace{-5.0mm}
\end{figure}

@@

\input{deformable-src.tex}

@@

\noindent{2) \em Multiscale Implementation.}
Performing alignment in a multiscale fashion has two benefits: first, it provides a larger region of attraction, and second, it reduces overall computational cost. Here, we further investigate the convergence behavior of the algorithm as a function of the standard deviation $\sigma$ of the Gaussian smoothing filter and the number of scales considered.
We use the same 7 illuminations in
Session 1 as training, and all 20 illuminations in the same
session as testing. We introduce artificial deformation in
both $x$ and $y$ directions up to 16 pixels in the
$80\times 60$ frame, with a step size of 4 pixels, i.e.,
$(\Delta x, \Delta y) \in \{-16,-12,\ldots,12,16\} \times
\{-16,-12,\ldots,12,16\}$. We consider an alignment
successful if the estimated coordinates of the eye-corners
are within 1 pixel from the ground truth in the original
image.  In Figure \ref{fig:multiscale}, we report the
alignment success rate, averaged over the artificially
perturbed initial deformations, as a function of the
standard deviation of the Gaussian kernel $\sigma$, for
three choices of the number of scales. As one can see,
using multiscale indeed improves the performance, and when
3 scales are used, a smaller convolution kernel can achieve
a similar performance compared to a much larger kernel when
only 2 scales are used.
\begin{figure}
\centering
\includegraphics{Figure_4}
\vspace{-2mm}
\caption{{\bf Multiscale alignment.} This figure shows the average success rate of alignment over all possible perturbations. A smaller blur kernel can be applied to achieve certain level of performance when more scales are used.}
\label{fig:multiscale}
\vspace{-2mm}
\end{figure}

@@

We have designed a system that can acquire frontal images of a subject while
simultaneously illuminating the subject from all directions above horizontal. A sketch of the
system is shown in Figure \ref{fig:system}: The illumination
system consists of four projectors that display various bright
patterns onto the three white walls in the corner of a dark
room.  The light reflects off of the walls and illuminates the
user's head indirectly.  After taking the frontal illuminations
we rotate the chair by 180 degrees and take pictures from the
opposite direction.  Having two cameras speeds the process
since only the chair needs to be moved in between frontal and
rear illuminations. Our projector-based system has several
advantages over flash-based illumination systems for face recognition:
\begin{itemize}
\item The illuminations can be modified in software, rather than hardware.
\item It is easy to capture many different illuminations quickly.
\item Good coverage and distant illumination can be achieved simultaneously.
\item There is no need to mount anything on the walls or construct a large dome.
\item The system can be assembled from off-the-shelf hardware.
\end{itemize}
\begin{figure}
\centering
\includegraphics{Figure_6}
\caption{{\bf Training acquisition system:} Four projectors and two cameras controlled by one computer.}
\vspace{-2em}
\label{fig:system}
\end{figure}
With our projector system, our choice of illuminations is
constrained only by the need to achieve a good
SNR,\footnote{Since illuminations with more pixels illuminated
will have a better SNR (provided they don't saturate), there is
an engineering tradeoff between the SNR and the number of
training images.} avoid saturation, and achieve a reasonably
short acquisition time.  Two simplifying assumptions that we
make are that every pixel is either turned fully on or off in
every illumination, and that the illuminated regions do not
overlap.

@@

We ran two experiments to guide our choice of illuminations for
our large-scale experiments:
\begin{enumerate}
\item {\em Coverage Experiment.} In the first experiment we
    attempt to determine what coverage of the sphere is
    required to achieve good interpolation for test images.
    The subject was illuminated by 100 (50 front, 50 back)
    illuminations arranged in concentric rings centered at
    the front camera.  Subsets of the training images were
    chosen, starting at the front camera and adding a ring
    at a time.  Each time a ring was added to the training
    illumination set, the average $\ell^1$ registration
    error (residual) for a set of test images taken under
    sunlight was computed and plotted in Figure
    \ref{fig:illumination-patterns}(a).  The more rings of
    training illuminations are added, the lower the
    representation error becomes, with diminishing returns.
\item {\em Granularity Experiment.} In the second
    experiment we attempt to determine how finely divided
    the illumination sphere should be.  At the first
    granularity level, the projectors  illuminate the
    covered area uniformly.  At each subsequent granularity
    level each illuminated cell is divided in two along its
    longer side but intensity doubled.  For each
    granularity level the average $\ell^1$ registration
    error is computed as in the coverage experiment and
    shown in Figure \ref{fig:illumination-sufficiency}(b).
    Again, diminishing returns are observed as more
    illuminations are added.
\end{enumerate}
\begin{figure}
\centering
\includegraphics{Figure_7}
\caption{{\bf Illumination Patterns.}   The cells are illuminated in sequence.  For rear illuminations the sequence is reversed.  In the chosen pattern's rear illumination, the cells 1-5 and 7-11 are omitted for a total of 38 illuminations. The four rectangular regions correspond to the four projectors.  }
\label{fig:illumination-patterns}
\vspace{-.75em}
\end{figure}

@@

\renewcommand{\tempscale}{0.45}
\begin{figure}
\centering
\includegraphics{Figure_12}
\caption{{\bf Representative examples of categories C1-C3}. One row for each category.}\label{fig:examples1-3}
\vspace{-.1in}
\end{figure}

>>>>

\begin{figure}
\centering
\includegraphics[width=3.5in]{Figure_2}
\caption{{\bf Comparing alignment of a subject wearing sunglasses by
$\ell^1$ and $\ell^2$ minimization.}
{\bf Top:} alignment result of minimizing $\|\e\|_1$; {\bf Bottom:}
result of minimizing $\|\e\|_2$. (a) {\em Green (dotted):} Initial face boundary
given by the face detector, {\em Red (solid):} Alignment result shown on the same
face; (b) warped testing image using the estimated transformation $\y_0$;
(c) reconstructed face $A_i\x$ using the training; (d) image of error $\e$. }\label{fig:L1-L2-align}
\vspace{-5.0mm}
\end{figure}

@@

\begin{algorithm}[thb]
\caption{\bf (Deformable Sparse Recovery and Classification for
Face Recognition)} \label{alg:deformable-src}
\begin{algorithmic}[1]
\begin{small}
\STATE {\bf Input:} Training images $\{A_i \in \Re^{m\times n_i}\}_{i=1}^K$ for $K$ subjects,  a test image
$\y\in\Re^m$ and a deformation group $T$.
\STATE
{\bf for} each subject $i$,
\STATE \hspace{3mm} $\tau^{(0)}
\leftarrow I$.
\STATE \hspace{3mm} {\bf while} not converged $(j=1,2,\ldots)$ {\bf do}
\STATE \hspace{6mm}
$\tilde \y(\tau) \leftarrow \frac{\y \circ \tau}{\|\y \circ
\tau\|_2}$; \;\;\; $J \leftarrow  \frac{\partial}{\partial
\tau} \tilde\y(\tau)  \bigr|_{\tau^{(j)}} $;
%\STATE \hspace{6mm} $(\hat \x, \hat \e, \Delta \tau) \leftarrow \left\{\begin{array}{l} \arg \min_{\x,\e,\Delta \tau} \| \e \|_1 \\  \subj \; \y + J \Delta \tau = A_k \x + \e \end{array}\right.$
\STATE \hspace{6mm} $ \Delta \tau =  \arg\min \; \| \e \|_1  \;
\subj \; \tilde \y + J \Delta \tau = A_i \x + \e.$
\STATE
\hspace{6mm} $\tau^{(j+1)} \leftarrow \tau^{(j)} + \Delta
\tau$;
\STATE \hspace{3mm} {\bf end while} \STATE {\bf end} \STATE Keep
the top $S$ candidates $k_1, \ldots, k_S$ with the smallest
residuals $\|\e\|_1$. \STATE Compute an average transformation
$\bar{\tau}$ from $\tau_{k_1}, \tau_{k_2}, \ldots, \tau_{k_S}$.
\STATE Update $\y \leftarrow \y \circ \bar{\tau}$ and $\tau_i
\leftarrow \tau_i \cdot \bar{\tau}^{-1}$ for $i = k_1, \dots,
k_S$. \STATE Set $A \leftarrow \big[ A_{k_1} \circ
\tau_{k_1}^{-1} \mid A_{k_2} \circ \tau_{k_2}^{-1} \mid \dots
\mid A_{k_S} \circ \tau_{k_S}^{-1} \big]$. \STATE Solve the
$\ell^1$-minimization problem: \hspace{2mm}
$\hat{\x} = \arg \min_{\x, \e} \| \x \|_1 + \|\e\|_1 \;\; \subj \;\; \y = A \x + \e.$
\STATE Compute residuals $r_i(\y) = \| {\y} - {A}_i \, \delta_i(\hat{\x}) \|_2$ for $i = k_1, \dots, k_S$.
\STATE {\bf Output:} $\mbox{identity}(\y) = \arg\min_i r_i(\y)$.
\end{small}
\end{algorithmic}
\end{algorithm}
%\vspace{-4mm}

@@

\noindent{2) \em Multiscale Implementation.}
Performing alignment in a multiscale fashion has two benefits: first, it provides a larger region of attraction, and second, it reduces overall computational cost. Here, we further investigate the convergence behavior of the algorithm as a function of the standard deviation $\sigma$ of the Gaussian smoothing filter and the number of scales considered.
We use the same 7 illuminations in
Session 1 as training, and all 20 illuminations in the same
session as testing. We introduce artificial deformation in
both $x$ and $y$ directions up to 16 pixels in the
$80\times 60$ frame, with a step size of 4 pixels, i.e.,
$(\Delta x, \Delta y) \in \{-16,-12,\ldots,12,16\} \times
\{-16,-12,\ldots,12,16\}$. We consider an alignment
successful if the estimated coordinates of the eye-corners
are within 1 pixel from the ground truth in the original
image.  In Figure \ref{fig:multiscale}, we report the
alignment success rate, averaged over the artificially
perturbed initial deformations, as a function of the
standard deviation of the Gaussian kernel $\sigma$, for
three choices of the number of scales. As one can see,
using multiscale indeed improves the performance, and when
3 scales are used, a smaller convolution kernel can achieve
a similar performance compared to a much larger kernel when
only 2 scales are used.
\begin{figure}
\centering
\includegraphics[width=3.4in]{Figure_4}
\vspace{-2mm}
\caption{{\bf Multiscale alignment.} This figure shows the average success rate of alignment over all possible perturbations. A smaller blur kernel can be applied to achieve certain level of performance when more scales are used.}
\label{fig:multiscale}
\vspace{-2mm}
\end{figure}

@@

We have designed a system that can acquire frontal images of a subject while
simultaneously illuminating the subject from all directions above horizontal. A sketch of the
system is shown in Figure \ref{fig:system}: The illumination
system consists of four projectors that display various bright
patterns onto the three white walls in the corner of a dark
room.  The light reflects off of the walls and illuminates the
user's head indirectly.  After taking the frontal illuminations
we rotate the chair by 180 degrees and take pictures from the
opposite direction.  Having two cameras speeds the process
since only the chair needs to be moved in between frontal and
rear illuminations. Our projector-based system has several
advantages over flash-based illumination systems for face recognition:
\begin{itemize}
\item The illuminations can be modified in software, rather than hardware.
\item It is easy to capture many different illuminations quickly.
\item Good coverage and distant illumination can be achieved simultaneously.
\item There is no need to mount anything on the walls or construct a large dome.
\item The system can be assembled from off-the-shelf hardware.
\end{itemize}
\begin{figure}
\centering
\includegraphics[width=2.9in]{Figure_6}
\caption{{\bf Training acquisition system:} Four projectors and two cameras controlled by one computer.}
\vspace{-2em}
\label{fig:system}
\end{figure}
With our projector system, our choice of illuminations is
constrained only by the need to achieve a good
SNR,\footnote{Since illuminations with more pixels illuminated
will have a better SNR (provided they don't saturate), there is
an engineering tradeoff between the SNR and the number of
training images.} avoid saturation, and achieve a reasonably
short acquisition time.  Two simplifying assumptions that we
make are that every pixel is either turned fully on or off in
every illumination, and that the illuminated regions do not
overlap.

@@

We ran two experiments to guide our choice of illuminations for
our large-scale experiments:
\begin{enumerate}
\item {\em Coverage Experiment.} In the first experiment we
    attempt to determine what coverage of the sphere is
    required to achieve good interpolation for test images.
    The subject was illuminated by 100 (50 front, 50 back)
    illuminations arranged in concentric rings centered at
    the front camera.  Subsets of the training images were
    chosen, starting at the front camera and adding a ring
    at a time.  Each time a ring was added to the training
    illumination set, the average $\ell^1$ registration
    error (residual) for a set of test images taken under
    sunlight was computed and plotted in Figure
    \ref{fig:illumination-patterns}(a).  The more rings of
    training illuminations are added, the lower the
    representation error becomes, with diminishing returns.
\item {\em Granularity Experiment.} In the second
    experiment we attempt to determine how finely divided
    the illumination sphere should be.  At the first
    granularity level, the projectors  illuminate the
    covered area uniformly.  At each subsequent granularity
    level each illuminated cell is divided in two along its
    longer side but intensity doubled.  For each
    granularity level the average $\ell^1$ registration
    error is computed as in the coverage experiment and
    shown in Figure \ref{fig:illumination-sufficiency}(b).
    Again, diminishing returns are observed as more
    illuminations are added.
\end{enumerate}
\begin{figure}
\centering
\includegraphics[width=3.1in]{Figure_7}
\caption{{\bf Illumination Patterns.}   The cells are illuminated in sequence.  For rear illuminations the sequence is reversed.  In the chosen pattern's rear illumination, the cells 1-5 and 7-11 are omitted for a total of 38 illuminations. The four rectangular regions correspond to the four projectors.  }
\label{fig:illumination-patterns}
\vspace{-.75em}
\end{figure}

@@

\begin{figure}
\centering
\includegraphics{Figure_12}
\caption{{\bf Representative examples of categories C1-C3}. One row for each category.}\label{fig:examples1-3}
\vspace{-.1in}
\end{figure}
